---
title: "Portfolio1: MLM"
author: "Yoo Ri Hwang"
date: "1/15/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Source 



https://stat.utexas.edu/images/SSC/documents/SoftwareTutorials/MultilevelModeling.pdf

https://cehs-research.github.io/eBook_multilevel/mlm-2-levels-pupil-popularity.html

I refer the code on these websites. 
I followed the general tutorials instructed on these website.

However, I made some changes. (I do some extra analysis that was not on the website etc. )
I also explained by myself. 

###   notification 
ICC calculated based on the null model. 

# Prepare

##  load library
```{r, warning=FALSE, message=FALSE}

library(foreign)
library(tidyverse)
library(haven)    
library(furniture)    
#install.packages('furniture')
library(stargazer)    
library(texreg)   
#install.packages('texreg')
library(gridExtra)    
#install.packages('gridExtra')
library(psych)        
library(car)      
#install.packages('car')
library(nlme)
library(lme4)         
library(lmerTest)     
library(optimx)
#install.packages('optimx')
# oh my gosh this never ends. 

# packages <- c('performance','interactions','HLMdiag','sjstats')
# 
# installed_packages <- packages %in% rownames(installed.packages())
# if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }


#devtools::install_github("goodekat/redres")

library(performance)  
library(interactions) 
library(HLMdiag)      
library(redres)       
library(sjstats)     
library(predict3d)

```

## load data
```{r, message=FALSE, warning=FALSE}
data_raw <- haven::read_sav("https://github.com/MultiLevelAnalysis/Datasets-third-edition-Multilevel-book/raw/master/chapter%202/popularity/SPSS/popular2.sav")

summary(data_raw)
```
data structure:

pupils are nested in the class. 
trying to predict popluarity of pupil by pupil's gender and extravert. Furthermore, teacher's experience (class-level variable) will be considered. 

To sum up, we have

Level-1 IV:
pupil gender (SEX): coded as boy=0, girl=1  
Pupil's extroversion (extrav) : 10-point scale.  

Level-2IV:
teacher's experience (texp)

subsetting the data

```{r, warning=FALSE}

# we only take what we need

data_pop <- data_raw %>%
  dplyr::mutate(id = paste(class,pupil, sep = "-") %>% #creating uniqe ids for student 
                  factor()) %>%
  dplyr::select(id,pupil:popteach)

tibble::glimpse(head(data_pop,10))
```

check our fancy data

```{r, warning=FALSE}

data_pop %>%
  dplyr::select(extrav, sex, texp, popular) %>%
  data.frame() %>%
  stargazer::stargazer(header=F,type="text")

```

## simple plots 

1. when we overlook nested structure 
```{r, warning=FALSE}
data_pop %>%
  ggplot() +
  aes(x = extrav, y = popular)+ 
  geom_point()+
  stat_smooth(method=lm)
```
2. consider cluster (group)

just for curious, we pick first 9 class and see 
```{r, warning=FALSE}
data_pop %>%
  dplyr::filter(class<=9)%>%
  ggplot() +
  aes(x = extrav, y = popular)+ 
  geom_point()+
  stat_smooth(method=lm)+
  facet_wrap(~ class) 

```
3.

```{r, message=F, warning=F}

ggplot(data      = data_pop,
       aes(x     = extrav,
           y     = popular,
           col   = class,
           group = class)) + 
  geom_point(size     = 1,
             alpha    = .5,
             position = "jitter") +
  theme_minimal() +
  theme(legend.position = "none") +
     geom_smooth(method = lm,
              se     = FALSE,
              size   = .5, 
              alpha  = .8)+ # to add regression line
  labs(title    = "Popularity vs. Extraversion")
  
  
```
This code (for third graph) was adpated from "https://www.rensvandeschoot.com/tutorials/lme4/"

## Centering

In the tutorial website, the author did not do the centering and keep analyzing. However, in most cases, it is not useful unless it has certian theoritical reasons. 

```{r}
#install.packages('misty')
library(misty)

#center level-1 predictor at the grandmean
data_pop$cgm_extrav<-misty::center(data_pop$extrav)
data_pop$cgm_sex<-misty::center(data_pop$sex)
data_pop$cgm_texp<-misty::center(data_pop$texp)

#center level-1 predictor at the clustermean
data_pop$cwc_extrav<-misty::center(data_pop$extrav, type="CWC", cluster = data_pop$class)
data_pop$cwc_sex<-misty::center(data_pop$sex, type="CWC", cluster = data_pop$class)

```
## Single-level equation (regression)


$$POP_{ij} = \beta_{0} + \beta_{1}SEX_{ij} + \beta_{2}EXT_{ij}+e_{ij}$$
Random parameter: $var[e_{ij}]$  
fixed: intercept, slopes.

I won't check the interaction at this point. 
This single-level regression is rediculous in that it ignore nonindependenc. 

```{r, warning=FALSE}

pop_lm<- lm(popular ~ sex + extrav, data = data_pop) 

summary(pop_lm)


```


# MULTILEVEL ANALYSIS 



## 1. intercept-only model (unconditional model, a.k.a. null model)

level 1:   
$$y_{ij} = \beta_{0j} + e_{ij}$$
level2:  
$\beta_{0j}$ is the intercept, and this can be expressed as:  

$$ \beta_{0j} = \gamma_{00} + u_{0j}$$
$\gamma_{00}$  =  (overall mean)   
$u_{0j}$ = class effect, the variance of the mean for each class around the overall mean of DV (popular) score. 

Combined together : mixed model 

$$y_{ij} = \gamma_{00} + u_{0j} + e_{ij}$$
random effect: var[$e_{ij}$], var[$u_{0j}$]
fixed effect:  $\gamma_{00}$


```{r, message=FALSE}

#My preference towards REML :D

null<-lmer(popular~1 + (1|class), data=data_pop, REML=TRUE)
summary(null)



```

Interpretaion: grandmean of pop of all students = 5.0779
class average tends to vary by 0.08738 (approximately)

ICC : proportion of between-class variance and total variance.

$$\rho = \frac{\sigma^{2}_{u0}}{\sigma^{2}_{u0} + \sigma^{2}_{e}}$$



By hand... 
$$\rho = \frac{0.7021}{0.7021 + 1.2218}$$



or..using a great package..

```{r,message=F, warning=F}
performance::icc(null)

```

interpretation:
36.5% of variance is due to the group


## 2.  one level-1 predictor (fixed)

level 1:   
$$y_{ij} = \beta_{0j} + \beta_{1j}Extra_{ij} + e_{ij}$$
level2:  

level2:
$$ \beta_{0j} = \gamma_{00} + u_{0j}$$

$$ \beta_{1j} = \gamma_{00}$$ 
we do not have a error term since this is fixed. 

Combined together : mixed model 

$$y_{ij} = \gamma_{00} + \gamma_{10}Extra_{ij} + u_{0j} + e_{ij}$$
random effect: var[$e_{ij}$], var[$u_{0j}$]  
fixed effect:  $\gamma_{00}$, $\gamma_{10}$


```{r, message=FALSE}

#My preference towards REML :D

onefix<-lmer(popular~1 + cgm_extrav + (1|class), data=data_pop, REML=TRUE)
summary(onefix)


```


## 3.  one level-1 predictor (random)
level 1:   
$$y_{ij} = \beta_{0j} + \beta_{1j}Extra_{ij} + e_{ij}$$
level2:  

level2:
$$ \beta_{0j} = \gamma_{00} + u_{0j}$$

$$ \beta_{1j} = \gamma_{10} + u_{1j}$$ 


Combined together : mixed model 

$$y_{ij} = \gamma_{00} + \gamma_{10}Extra_{ij} + u_{0j} + u_{1j}Extrav_{ij} + e_{ij}$$


random effect: var[$e_{ij}$], var[$u_{0j}$]  
fixed effect:  $\gamma_{00}$, $\gamma_{10}$


```{r, message=FALSE}

#My preference towards REML :D

oneram<-lmer(popular~1 + cgm_extrav + (1 + cgm_extrav|class), data=data_pop, REML=TRUE)
summary(oneram)


```







##  4. two level-1 predictor (random)

level 1:   
$$y_{ij} = \beta_{0j} + \beta_{1j}EXTRA_{ij} + \beta_{1j}SEX_{ij} + e_{ij}$$
level2:  

level2:
$$ \beta_{0j} = \gamma_{00} + u_{0j}$$

$$ \beta_{1j} = \gamma_{10} + u_{1j}$$ 
$$ \beta_{2j} = \gamma_{20} + u_{2j}$$ 


Combined together : mixed model 

$$y_{ij} = \gamma_{00} + \gamma_{10}Extra_{ij} + \gamma_{20}SEX_{ij} + u_{0j} + u_{1j}Extrav_{ij} + u_{2j}SEX_{ij} + e_{ij}$$




```{r}



tworam<-lmer(popular~1 + cgm_extrav + cgm_sex + (1 + cgm_extrav+cgm_sex|class), data=data_pop)
summary(tworam)


```
model failed (convergency problem)
with a random effect with a variance very close to zero, this model should be reconsidered : convergency  

Alternatively, using ML: 
```{r}


tworam<-lmer(popular~1 + cgm_extrav + cgm_sex + (1 + cgm_extrav+cgm_sex|class), data=data_pop, REML=FALSE)
summary(tworam)
isSingular(tworam, tol = 1e-4)
```
 matrix argument is singular.
 
 
##  5. one level-2 predictor and two level-1 predictor (w/o interaction)


level 1:   
$$y_{ij} = \beta_{0j} + \beta_{1j}EXTRA_{ij} + \beta_{2j}SEX_{ij} + e_{ij}$$

level2:
$$ \beta_{0j} = \gamma_{00} +\gamma_{01}Texp_{j} + u_{0j}$$

$$ \beta_{1j} = \gamma_{10} + u_{1j}$$ 
$$ \beta_{2j} = \gamma_{20} + u_{2j}$$ 


Combined together : mixed model 

$$y_{ij} = \gamma_{00} +\gamma_{01}Texp_{j} + \gamma_{20}SEX_{ij} +  u_{1j}Extrav_{ij} + u_{2j}SEX_{ij} + u_{0j}+e_{ij}$$



```{r, message=FALSE}

#My preference towards REML :D
# 
inter<-lmer(popular~1+cgm_extrav+cgm_sex+texp + (1+cgm_extrav+cgm_sex|class), data = data_pop)
summary(inter)


```


## 6. two levle-1 predictor, two level 1 predictior (with cross-level interaction)


level 1:   
$$y_{ij} = \beta_{0j} + \beta_{1j}EXTRA_{ij} + \beta_{2j}SEX_{ij} + e_{ij}$$

level2:
$$ \beta_{0j} = \gamma_{00} +\gamma_{01}Texp_{j} + u_{0j}$$

$$ \beta_{1j} = \gamma_{10} +\gamma_{11}Texp_{j}+ u_{1j}$$ 
$$ \beta_{2j} = \gamma_{20} +\gamma_{21}Texp_{j}+ u_{2j}$$ 


Combined together : mixed model 

$$y_{ij} = \gamma_{00} +\gamma_{01}Texp_{j} + \gamma_{10}Extra_{ij} + \gamma_{20}SEX_{ij} + \gamma_{11}Texp_{j}*Extrav_{ij} + \gamma_{21}Texp_{j}*Sex_{ij} + u_{1j}*Extrav_{ij} + u_{2j}SEX_{ij} + u_{0j}+e_{ij}$$


```{r}

#My preference towards REML :D
# 
inter<-lmer(popular~1+cgm_extrav+cgm_sex+cgm_texp + cgm_extrav*cgm_texp+cgm_sex*cgm_texp+ (1+cgm_extrav+cgm_sex|class), data = data_pop)
summary(inter)
isSingular(inter, tol = 1e-4)
# the matrix argument is singular. 
```
If i did not center level 2 variable, it is like this. 
```{r}

inter<-lmer(popular~1+cgm_extrav+cgm_sex+texp + cgm_extrav*texp+cgm_sex*texp+ (1+cgm_extrav+cgm_sex|class), data = data_pop)
summary(inter)

```



