---
title: "Untitled"
author: "Yoo Ri Hwang"
date: "3/28/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this project, I tried SEM.. like licking the surface of the SEM.

.

Also, I refered many materials (see the readme file).



### What is SEM
Source: in the readme file. 

Structure Equation Modeling (SEM) analysis the network of relationship between the variables (including latent constructs, and measured variable ).

Unlike other conventional stat techniques, SES reqiures model specification( including measurement error specification)

SES is linear statistics technique, and do not provide causality test.

Keep in mind that SES do not provide default model. 
SEM has a powerful weapon- intuitive visulaization.

*Goals*
1) patterns of cov/cor among the variables
2) how much variance can be explianed by this specificed model. 


### Packages


```{r}


library(lavaan)
library(tidyverse)
library(psych)
library(MASS)
library(mvnormalTest)
library(semPlot)

```


### Simulate the data

```{r}
demo.model <- '
y ~ .5*f 

f =~ .8*x1 + .8*x2 + .8*x3 + .8*x4 + .8*x5  

x1 ~~ (1-.8^2)*x1
x2 ~~ (1-.8^2)*x2
x3 ~~ (1-.8^2)*x3
x4 ~~ (1-.8^2)*x4
x5 ~~ (1-.8^2)*x5
'
simData <- lavaan::simulateData(demo.model, sample.nobs=200)

# see the describtive stats
psych::describe(simData)

# multivariate nurmality test 

mv<-mardia(simData)
mv

#?simulateData

sim.cor <-cor(simData, use="pairwise.complete.obs", method="pearson")
sim.cor

```
### specify the model 
```{r}


tofit.model <- '
y ~ f 
f =~ x1+ x2 + x3 + x4 + x5 
x1 ~~ x1 
x2 ~~ x2 
x3~~x3 
x4~~x4 
x5~~x5 
'

tofit.model_m <- sem(tofit.model, simData)

summary(tofit.model_m, fit.measures=T)
inspect(tofit.model_m)
semPlot::semPaths(tofit.model_m)
```


lavaan package's notation

**~ : predict** 
(y~x: y is predicted by x)

**=~ : indicator**
(latent variable =~ observed indicator)

**~~** 
Covariance


**~1: intercept or mean **
x~1 : mean of variable x

**1*** fixed parameter or loading to one

**NA*** frees parameter or loading 

**a***labels the parameter 'a'
for model constraints 

## Simple regression 

before going further, let's try simple regression

```{r}

hey<-lm(y~x2, simData)
summary(hey)

```

results interpretaion:

as one unit increase in x1, the y score improves approaximatly 0.508, and Significant. 
the residual standard error is 1.087. 
the residual variance is (1.087)^2 = 1.181569

This can be done by Lavaan packages

```{r}
hey2<-'
y ~ 1 + x2
x2~~x2
'

hey2fit<-sem(hey2,data=simData)
summary(hey2fit)
semPlot::semPaths(hey2fit)

```

results interpretaion:

the regression coifficient (0.508) is the same.

However, some difference is here. ML and OLS provide different residual variance but the same coefficients. 


## Path analysis

What is different between SEM and path analysis?

"SEM is a combination of multiple regression and factor analysis. Path analysis deals only with measured variables. two or more measured variables"
(Source:https://theicph.com/wp-content/uploads/2016/09/How-to-conduct-Path-Analysis-and-SEM-for-Health-Research_24-Sep-2016_Prof-Bhisma-Murti.pdf)
####  Data simulation 
```{r}

set.seed(1234)
x<-rnorm(100)
m<-0.5*x + rnorm(100)
y<-0.7*m + rnorm(100)
data <-data.frame(x=x,m=m,y=y)



```

#### model specification 

```{r}


medmodel<-'
y~c*x
m~a*x
y~b*m
# indirect effect (a*b)
ab:=a*b
# total effect
total:=c+(a*b)
'

```


```{r}
medmodel_m<-sem(medmodel, data=data)
summary(medmodel_m, fit.measures=T)
semPaths(medmodel_m)
```

## Confirmatory factor analysls
load the data 
```{r}
library(foreign)
dat <- read.spss("https://stats.idre.ucla.edu/wp-content/uploads/2018/05/SAQ.sav",
                                             to.data.frame=TRUE, use.value.labels = FALSE)
head(dat,10)


```
8 items, with one factor 
```{r}

cfa<- '
f =~ q01 + q02 + q03 + q04 + q05 + q06 + q07 + q08'

cfa8<- cfa(cfa, data=dat, std.lv=TRUE)
#std.lv=T automatically standardize the variance. 

summary(cfa8, fit.measures=T, standardized=T)
semPaths(cfa8)
```

interpretaion 


```{r}
round(cor(dat[,1:8]),2)


```
q 02, it is only  -.23.
If we see the corr table, also can find that q02 is weakly associated with other questions. 
q03, 